---
title: "Homework 4 Answers"
subtitle: "BSTA 512/612"
author: "Your name here!!!"
description: "Due: Friday February 28, 2025 at 11pm"
date-modified: "today"
format: 
  html:
    link-external-newwindow: true
    toc: true
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

Answers are not necessarily complete! This is just meant to serve as a check if you are stuck. 

```{=html}
<style>div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 4, fig.width = 6, fig.align = "center")

library(tidyverse)
library(rstatix)
library(broom)
library(gt)
library(janitor)
library(readxl)
library(gridExtra)
library(ggfortify)  # autoplot(model)
library(gtsummary)
library(describedata) # gladder()
library(haven)
library(here)
```

## Questions

### Question 1

```{r}
#| echo: false
dep_df = read_sas(here("data/completedata.sas7bdat"))
```

#### Part a

Fit the regression model with all the covariates (Fatalism, Optimism, Spirituality), display the regression table, and write out the fitted regression line.

::: blue
```{r}
#| echo: false


q2_mod_f1 = lm(Depression ~ Fatalism + Optimism + Spirituality, 
                data = dep_df)

q2_mod_ft = tidy(q2_mod_f1, conf.int = T)

tbl_regression(q2_mod_f1, intercept = T)
```
$$\begin{aligned}
\widehat{\text{Depression}} &= `r round(q2_mod_ft$estimate[1], 1)` + `r round(q2_mod_ft$estimate[2], 2)` \cdot \text{Fatalism} `r round(q2_mod_ft$estimate[3], 2)` \cdot \text{Optimism} + `r round(q2_mod_ft$estimate[4], 2)` \cdot \text{Spirituality}
\end{aligned}$$
:::

#### Part b

Does at least one of the covariates contribute significantly to the prediction of Depression? (Note: this is an overall test. Please follow the hypothesis test steps. To complete step 4-6, simply output your ANOVA table.)

::: blue
```{r}
#| echo: false

q2_mod_f1 = lm(Depression ~ Fatalism + Optimism + Spirituality, 
            data = dep_df) # already ran in earlier part
q2_mod_r1 = lm(Depression ~ 1, 
            data = dep_df) 
anova(q2_mod_r1, q2_mod_f1) %>% tidy() %>% gt() %>% fmt_number(decimals = 4)
```
:::

#### Part c

Does the addition of Spirituality add significantly to the prediction of Depression achieved by Fatalism and Optimism?

::: blue
```{r}
#| echo: false
q2_mod_f2 = lm(Depression ~ Fatalism + Optimism + Spirituality, 
            data = dep_df) # already ran in earlier part
q2_mod_r2 = lm(Depression ~ Fatalism + Optimism, 
            data = dep_df) 
anova(q2_mod_r2, q2_mod_f2) %>% tidy() %>% gt() %>% fmt_number(decimals = 2)
```
:::

#### Part d

::: blue
```{r}
#| echo: false
q2_mod_f3 = lm(Depression ~ Fatalism + Optimism + Spirituality, 
            data = dep_df) # already ran in earlier part
q2_mod_r3 = lm(Depression ~ Fatalism, 
            data = dep_df) 
anova(q2_mod_r3, q2_mod_f3) %>% tidy() %>% gt() %>% fmt_number(decimals = 2)
```
:::


### Question 2

```{r}
#| echo: false
protein <- read_excel(here("data/CH12Q03.xls")) # change this to the appropriate file location
```

#### Part a

Using R, make a variable that is a factor for Diet. Make sure to check what values the original variable for Diet can take. How many indicator functions do you need to represent the categorical variable Diet (protein-rich vs. protein-poor)?

::: blue
```{r}
#| include: false
protein = protein %>%
  mutate(DIET = factor(DIET, levels = c("POOR", "RICH")))
```

2 levels, 1 indicator
:::

#### Part b

At a level of significance $\alpha = 0.10$, test whether protein diet modifies the effect of age on height. Justify your answer (e.g., perform a hypothesis test for the interaction between diet and age).

::: blue

```{r}
#| echo: false
mod_diet_red = lm(HT ~ AGE + DIET, data = protein)
mod_diet_full = lm(HT ~ AGE + DIET + AGE * DIET, data = protein)
```

```{r}
#| echo: false
anova(mod_diet_red, mod_diet_full) %>% tidy() %>% gt() %>% fmt_number(decimals = 4)
```
:::

#### Part c

Is it possible that diet is a confounder? Note: this will depend on your results from Part b. 

::: blue
:::

#### Part d

Write the fitted regression equation for our model in Part b. Write the respective regression lines for each specific diet group: protein rich and protein poor. Interpret the slope of each regression line (no need for a 95% CI here).

::: blue
**Fitted regression equation:**

$$\widehat{\text{Height}}|\text{Age}, \text{Diet} = 51.23+8.69\cdot \text{Age}-0.90\cdot I(\text{Protein-rich})+7.32\cdot \text{Age}\cdot I(\text{Protein-rich})$$
**Protein rich group $I(\text{Protein-rich})=1$:**

$$\widehat{\text{Height}}|\text{Age}, \text{Diet} = ??$$

**Protein poor group $I(\text{Protein-rich})=0$:**

$$\widehat{\text{Height}}|\text{Age}, \text{Diet} = ??$$

 

**Interpretation** You need to work through this!
:::

### Question 3

```{r}
#| echo: false

HDL_factors <- read_excel(here("data/CH09Q05.xls")) # change this to the appropriate file location
```

#### Part a

Use $\alpha= 0.05$, test whether if there is a crude association between HDL measurement and total cholesterol. Note: testing for a crude association means we fit a simple linear regression model and see if the association is significant. 

::: blue
```{r}
#| echo: false
model_X1 = lm(Y ~ X1, data = HDL_factors)
```

```{r}
#| echo: false
anova(model_X1) %>% tidy() %>% gt() %>% fmt_number(decimals = 3)
```

:::

#### Part b

Sometimes simple linear regression leads us to believe that there is no association between two variables, but missing interaction might be obscuring the association. Use $\alpha= 0.1$ to test whether total triglyceride is an effect modifier of the association between HDL and total cholesterol. 

::: blue
```{r}
#| echo: false
mod_X1_X2_full = lm(Y ~ X1 + X2 + X1*X2, data = HDL_factors)
mod_X1_X2_red = lm(Y ~ X1 + X2, data = HDL_factors)

anova(mod_X1_X2_red, mod_X1_X2_full) %>% tidy() %>% 
  gt() %>% fmt_number(decimals = 3)
```
:::


#### Part c

Is it possible that total triglyceride is a confounder? No need to test this explicity.

::: blue
:::


