---
title: "Homework 3 Answers"
subtitle: "BSTA 512/612"
author: "Your name here!!!"
description: "Due: Friday February 14, 2025 at 11pm"
date-modified: "today"
format: 
  html:
    link-external-newwindow: true
    toc: true
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

Answers are not necessarily complete! This is just meant to serve as a check if you are stuck. 

```{=html}
<style>div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}</style>
```

```{r setup, include=FALSE}
library(tidyverse)
library(rstatix)
library(broom)
library(gt)
library(janitor)
library(readxl)
library(gridExtra)
library(ggfortify)  # autoplot(model)
library(gtsummary)
library(describedata) # gladder()
library(haven)
library(here)
```

## Questions

### Question 1

This question and data are adapted from [this textbook](https://catalog.lib.uchicago.edu/vufind/Record/7546272).

In an experiment designed to describe the dose–response curve for vitamin K, individual rats were depleted of their vitamin K reserves and then fed dried liver for 4 days at different dosage levels. The response of each rat was measured as the concentration of a clotting agent needed to clot a sample of its blood in 3 minutes. The results of the experiment on 12 rats are given in the following table; values are expressed in common logarithms for both dose and response.

```{r}
clot = read_excel(here("data/CH05Q09.xls"))
clot %>% gt() %>%
  cols_label(RAT = md("**Rat**"),
             LOGCONC = md("**Log10 Concentration (Y)**"),
             LOGDOSE = md("**Log10 Dose (X)**"))
```

Use the log-transformed values as given in the dataset.

Use the following scatterplot to build your answers off of:

```{r}
#| fig-height: 5
#| fig-width: 6.5
#| fig-align: center
#| echo: false

ggplot(clot, aes(x = LOGDOSE,
                 y = LOGCONC)) +
  geom_point(size = 2) +
  labs(x = "Log10 dose (X)", 
       y = "Log10 concentration (Y)",
       title = "Concentration of clotting agent needed vs. dosage levels of liver") +
    theme(axis.title = element_text(size = 12), 
        axis.text = element_text(size = 12), 
        title = element_text(size = 12))
```

#### Part a

::: blue
Here is the code for fitting the model:

```{r}
#| echo: false 
clot_mod <- lm(LOGCONC ~ LOGDOSE, data = clot)
```

```{r}
summary(clot_mod)
```
:::

#### Part b

::: blue
```{r}
#| echo: false
anova(clot_mod) %>% tidy() %>% gt()
```
:::

#### Part c

::: blue
$$ F = 1148.759 $$
::: 

#### Part d

::: blue
Not shown bc answer is complete solution
:::

### Question 2

#### Part a

::: blue
```{r}
#| echo: false
(r <- cor(clot$LOGCONC, clot$LOGDOSE))
```
:::

#### Part b

::: blue
```{r}
#| echo: false
anova(clot_mod) %>% tidy() %>% gt()

clot_anova = anova(clot_mod) %>% tidy()

SSY = sum(clot_anova$sumsq)
SSR = clot_anova$sumsq[1]

(R2 = SSR/SSY)
```
:::

#### Part c

::: blue
Not given
:::


### Question 3

```{r}
#| message: false

if(!require(Sleuth3)) { install.packages("Sleuth3"); library(Sleuth3) }
q1_data = ex0824 
```

#### Part a

::: blue
```{r}
#| fig-width: 6
#| fig-height: 4.5
#| fig-align: center
#| echo: false
#| warning: false

ggplot(data = q1_data, aes(x = Age, y = Rate)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth(se = F, color = "#F14124") +
  labs(title = "Scatterplot of Respiratory Rate vs. Age", 
       x = "Age (months)", 
       y = "Respiratory Rate (breaths/minute)")
```
:::

#### Part b

::: blue
Not given
:::

#### Part c

Fit the regression model, display the regression table, and write out the fitted regression line.

::: blue
```{r}
#| echo: false
q1_model = lm (Rate ~ Age, data = q1_data)
q1_mod_t = tidy(q1_model)
q1_mod_t %>% gt()
```

$$\begin{aligned}
\widehat{\text{RR}} &= `r round(q1_mod_t$estimate[1], 2)` `r round(q1_mod_t$estimate[2], 2)` \cdot \text{Age}
\end{aligned}$$
:::

#### Part d

::: blue
```{r}
#| echo: false
# augment is from the broom package
q1_aug <- augment(q1_model)
names(q1_aug) # Just me looking at the names to call in each plot
```

```{r fig.height=3, fig.width=8}
#| echo: false
hist1 <- ggplot(q1_aug, aes(x = .resid)) +
  geom_histogram()

density1 <- ggplot(q1_aug, aes(x = .resid)) +
  geom_density()

box1 <- ggplot(q1_aug, aes(x = .resid)) +
  geom_boxplot()

grid.arrange(hist1, density1, box1, nrow = 1)
```
:::

#### Part e

::: blue

```{r fig.width=3, fig.height=3}
#| echo: false
ggplot(q1_aug, aes(sample = .resid)) + 
  stat_qq() +     # points
  stat_qq_line() +   # line
  labs(y = "Fitted residuals", 
       x = "Theoretical quantiles")
```
:::

#### Part f

::: blue
```{r}
#| echo: false
shapiro.test(q1_aug$.resid) %>% tidy() %>% gt()
```
:::

#### Part g

::: blue
```{r}
#| echo: false
ggplot(q1_aug, aes(x = Age, y = .resid)) + # may also use y = .std.resid
  geom_point() +
  geom_hline(yintercept = 0, color = "#FF8021") +
  labs(title = "Residual plot")
```
:::

#### Part h

::: blue
```{r}
#| echo: false
n = nobs(q1_model)
q1_aug %>% filter(.hat > 6/n) %>%
  arrange(desc(.hat))
```
:::

#### Part i

::: blue
```{r}
#| echo: false
q1_aug %>% arrange(desc(.cooksd))
```
:::

#### Part j

::: blue
```{r}
#| echo: false
ggplot(q1_data, aes(x = Rate)) + geom_histogram()
```
:::

#### Part k

::: blue

```{r}
#| echo: false
gladder(q1_data$Rate)
```
:::

#### Part l

::: blue
Not given

```{r}
#| echo: false
q1_data2 = q1_data %>%
  mutate(Rate_log = log(Rate), 
         Rate_sr = sqrt(Rate), 
         Rate_inv_sr = 1/sqrt(Rate))
```
:::

#### Part m

::: blue
```{r}
#| fig-width: 6
#| fig-height: 4.5
#| fig-align: center
#| message: false
#| echo: false

ggplot(data = q1_data2, aes(x = Age, y = Rate_log)) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method = "lm", color = "darkgreen") +
  labs(title = "Scatterplot of Log Respiratory Rate vs. Age", 
       x = "Age (months)", 
       y = "Log Respiratory Rate (log breaths/minute)")

ggplot(data = q1_data2, aes(x = Age, y = Rate_sr)) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method = "lm", color = "darkgreen") +
  labs(title = "Scatterplot of Square root Respiratory Rate vs. Age", 
       x = "Age (months)", 
       y = "SqRoot Respiratory Rate \n (square root breaths/minute)")

ggplot(data = q1_data2, aes(x = Age, y = Rate_inv_sr)) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method = "lm", color = "darkgreen") +
  labs(title = "Scatterplot of Inverse square root Respiratory Rate vs. Age", 
       x = "Age (months)", 
       y = "Inverse SqRoot Respiratory Rate \n (inverse square root breaths/minute)")
```
:::

#### Part n

::: blue

**For the log-transformed Rate:**
```{r}
q1_model2 = lm (Rate_log ~ Age, data = q1_data2)
q1_mod_t2 = tidy(q1_model2)
q1_mod_t2 %>% gt()
```
:::

#### Part o

::: blue
**For log-transformed Rate:**
```{r fig.width=3, fig.height=3}
#| echo: false
q1_aug2 <- augment(q1_model2)
ggplot(q1_aug2, aes(sample = .resid)) + 
  stat_qq() +     # points
  stat_qq_line() +   # line
  labs(y = "Fitted residuals", 
       x = "Theoretical quantiles")
```
:::


#### Part p

::: blue

**For log-transformed Rate:**
```{r fig.width=4, fig.height=3}
#| echo: false
ggplot(q1_aug2, aes(x = Age, y = .resid)) + # may also use y = .std.resid
  geom_point() +
  geom_hline(yintercept = 0, color = "#FF8021") +
  labs(title = "Residual plot")
```
:::


#### Part q

::: blue
Not given
:::

### Question 4

```{r}
dep_df = read_sas(here("data/completedata.sas7bdat"))
```

#### Part a

::: blue

```{r}
#| echo: false
q2_mod_f1 = lm(Depression ~ Fatalism + Optimism + Spirituality, 
                data = dep_df)

q2_mod_ft = tidy(q2_mod_f1, conf.int = T)

q2_mod_ft %>% gt() %>%
  tab_options(table.font.size = 15) %>% fmt_number(decimals = 4)
```

Another fun way to display: 
```{r}
tbl_regression(q2_mod_f1, intercept = T)
```
:::

#### Part b

::: blue
-   $\beta_0$: The expected depression score is `r round(q2_mod_ft$estimate[1], 1)` when fatalism, depression, and spirituality scores are 0 (95% CI: `r round(q2_mod_ft$conf.low[1], 1)`, `r round(q2_mod_ft$conf.high[1], 1)`).

    -   Same as homework 2: The intercept does not make sense. A score of 0 is outside the range of possible scores for fatalism, optimism, and spirituality.

-   $\beta_1$: For every 1 point higher fatalism score, there is an expected difference of `r round(q2_mod_ft$estimate[2], 2)` points higher depression score, adjusting for optimism and spirituality score (95% CI: `r round(q2_mod_ft$conf.low[2], 2)`, `r round(q2_mod_ft$conf.high[2], 2)`).
:::

#### Part c

::: blue
Not given
:::

#### Part d

::: blue
$$\begin{aligned}
\widehat{\text{Depression}} &= `r round(q2_mod_ft$estimate[1]  + q2_mod_ft$estimate[4] * 6 + q2_mod_ft$estimate[3] * 10, 2)` + `r round(q2_mod_ft$estimate[2], 2)` \cdot \text{Fatalism}
\end{aligned}$$
:::